---
title: "p8105_hw2_yq2251"
author: "TritonD"
date: "9/27/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
library(readxl)
library(dplyr)
```

#import data sheet Mr. Trash Wheel
```{r}
MrTrashWheel_data = read_excel( "./HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet= "Mr. Trash Wheel", range = cell_cols("A:N"))%>%
 janitor::clean_names()%>%
drop_na(dumpster)%>%
  mutate(sports_balls=as.integer(round(sports_balls)))
```

# import datasheet Rain_2017 and Rain_2018
```{r}
Rain_2017 = read_excel( "./HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet= "2017 Precipitation", range = "A2:B14")%>%
janitor::clean_names()%>%
  drop_na(total)%>%
  mutate(year=2017)
```

```{r}
Rain_2018 = read_excel( "./HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet= "2018 Precipitation",range = "A2:B14")%>%
janitor::clean_names()%>%
  drop_na(total)%>%
  mutate(year=2018)
```

#Using full_join to combine 2017-18 data and rename "month"
```{r}
Rain_Total=
full_join(Rain_2017,Rain_2018, by=NULL)%>%
mutate(month=month.name[month])
```

#In dataset "MrTrashWheel_data", there are 14 columns with 285 observations. Key variables including "weight_tons" which shows the amount of garbage they were collected each day, "dumpster" represents the number of dumpsters collected.
#In dataset "Rain_Total", there are 3 columns with 19 observations.Key variables including "total" which represents the amount of precipitation in that month.
#Total precipitation in 2018 is 23.50 inches, the median number of sports balls in a dumpster in 2017 is `r median(MrTrashWheel_data$sports_balls, na.rm=FALSE)`.


#Question 2
#use separate function to separate "mon" and create variable "president"
```{r}
Pols = read_csv( "./pols-month.csv")%>%
  separate(mon, into = c("year", "month","day"), sep = "-")%>%
  mutate(month=as.integer(month),year=as.integer(year),day=as.integer(day))%>%
  mutate(month=month.name[month])%>%
  mutate(president= ifelse(prez_gop %in% 0, "dem",
        ifelse(prez_gop %in%1, "gop","gop")))%>%
  select(-prez_dem,-prez_gop,-day)
```
```{r}
snp=read_csv( "./snp.csv")%>%
  separate(date, into = c("day", "month","year"), sep = "/")%>%
  mutate(month=as.integer(month),year=as.integer(year),day=as.integer(day))%>%
  mutate(month=month.name[month])%>%
  select("year", everything())%>% 
  select(-day)
```
#use dataset unemployment,change abbreviated month name to full
```{r}
unemploy=read_csv( "./unemployment.csv")%>%
  rename(January=Jan, February=Feb, March=Mar, April=Apr,June=Jun,July=Jul,August=Aug,September=Sep,October=Oct,November=Nov, December=Dec)
```

```{r}
unemploy_tidy=
  pivot_longer(
    unemploy, 
    January:December,
    names_to = "month", 
    values_to = "unemployment_rate")
```

#join the datasets
```{r}
snp_pols=left_join(Pols,snp,by=NULL)
```
```{r}
snp_pols_unemploy=left_join(snp_pols,unemploy_tidy, by=NULL)
```
#Dataset "snp" contains 787 observations in 3 columns presenting Standard & Poor’s stock market index (S&P) from March 1950-Jan 2015.

#Dataset "Pols" has 822 observations of 9 variables presenting # of national politicians who are democratic or republican ranging from Jan1947-June 2015.

#Dataset "unemploy_tidy" has 816 observations of 3 columns presenting the unemployment rate from Jan 1948-June 2015.Important variable include "president" and "close".

#Dataset "snp_pols"has 1391 observations in 10 columns, which is a combination of dataset "snp" and "Pols".Important variable include "president", "close" and "unemployment_rate

#Dataset "snp_pols_unemploy" has 94,588 observations with 12 columns, which is a combination of dataset "snp_pols" and "unemploy_tidy"













