---
title: "p8105_hw2_yq2251"
author: "TritonD"
date: "9/27/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
library(readxl)
library(dplyr)
```

#import data sheet Mr. Trash Wheel
```{r}
MrTrashWheel_data = read_excel( "./HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet= "Mr. Trash Wheel", range = cell_cols("A:N"))%>%
 janitor::clean_names()%>%
drop_na(dumpster)%>%
  mutate(sports_balls=as.integer(round(sports_balls)))
```

# import datasheet Rain_2017 and Rain_2018
```{r}
Rain_2017 = read_excel( "./HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet= "2017 Precipitation", range = "A2:B14")%>%
janitor::clean_names()%>%
  drop_na(total)%>%
  mutate(year=2017)
```

```{r}
Rain_2018 = read_excel( "./HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet= "2018 Precipitation",range = "A2:B14")%>%
janitor::clean_names()%>%
  drop_na(total)%>%
  mutate(year=2018)
```

#Using full_join to combine 2017-18 data and rename "month"
```{r}
Rain_Total=
full_join(Rain_2017,Rain_2018, by=NULL)%>%
mutate(month=month.name[month])
```

#In dataset "MrTrashWheel_data", there are 14 columns with 285 observations. Key variables including "weight_tons" which shows the amount of garbage they were collected each day, "dumpster" represents the number of dumpsters collected.
#In dataset "Rain_Total", there are 3 columns with 19 observations.Key variables including "total" which represents the amount of precipitation in that month.
#Total precipitation in 2018 is 23.50 inches, the median number of sports balls in a dumpster in 2017 is `r median(MrTrashWheel_data$sports_balls, na.rm=FALSE)`.


#Question 2
#use separate function to separate "mon"
```{r}
Pols = read_csv( "./pols-month.csv")%>%
  separate(mon, into = c("year", "month","day"), sep = "-")%>%
  mutate(month=as.integer(month),year=as.integer(year),day=as.integer(day))%>%
  mutate(month=month.name[month])%>%
  mutate(president= ifelse(prez_gop %in% 0, "dem",
        ifelse(prez_gop %in%1, "gop","gop")))%>%
  select(-prez_dem,-prez_gop,-day)
```












